---
title: "Scraping_steam_V1_20260302"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Scraping Steam

## Setup

### Libraries

```{r}
# Scraping
library(selenium)
library(scrapex)
library(xml2)
library(magrittr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(stringr)
library(lubridate)
library(DataExplorer) # plot_intro, plot_missing
# API
library(httr2)
library(jsonlite)
library(ggplot2)
```

### Scrape function
This is a function that takes ANY steam search URL (hopefully, after much trial and error!) and returns the scraped data.
```{r}

harvest_steam_page <- function(url, target_games = 500) {
  
  # Navigate to the provided URL [cite: 7]
  driver$navigate(url)
  Sys.sleep(3)
  
  current_game_count <- 0
  previous_game_count <- -1
  
  # Scroll loop
  while(current_game_count < target_games && current_game_count > previous_game_count) {
    previous_game_count <- current_game_count
    driver$execute_script("window.scrollTo(0, document.body.scrollHeight);")
    Sys.sleep(2) 
    
    page_source <- driver$get_page_source() 
    
    current_game_count <- page_source |>
      read_html() |>
      xml_find_all("//a[contains(@class, 'search_result_row')]") |>
      length()
  }
  
  # Parse the HTML
  html_code <- page_source |> read_html()
  game_rows <- html_code |> xml_find_all("//a[contains(@class, 'search_result_row')]")
  
  safe_extract <- function(node, xpath) {
    element <- xml_find_first(node, xpath)
    if (is.na(element)) return(NA_character_)
    element |> xml_text() |> trimws()
  }
  
  # Build and return the tibble
  tibble(
    title = sapply(game_rows, safe_extract, ".//span[@class='title']"),
    release_date = sapply(game_rows, safe_extract, ".//div[contains(@class, 'search_released')]"),
    discount_pct = sapply(game_rows, safe_extract, ".//div[contains(@class, 'discount_pct')]"),
    original_price = sapply(game_rows, safe_extract, ".//div[contains(@class, 'discount_original_price')] | .//strike"),
    current_price = sapply(game_rows, safe_extract, ".//div[contains(@class, 'discount_final_price')]")
  )
}

```

### Clean function
Cleaning Function - should take a raw steam dataframe and format the numeric columns

```{r}
clean_steam_prices <- function(df) {
  df |>
    mutate(
      # dmy() parses the text and strictly returns a Date object
      release_date = dmy(release_date),
      # Remove '-' and '%' from discount, convert to numeric
      discount_pct = as.numeric(str_replace_all(discount_pct, "[-%]", "")),
      
      # Remove '€', replace ',' with '.', convert to numeric
      original_price = as.numeric(str_replace_all(original_price, "[€]", "") |> str_replace(",", ".")),
      current_price = as.numeric(str_replace_all(current_price, "[€]", "") |> str_replace(",", "."))
    )
}

# 2. Scalable Imputation Function - should fix missing data
impute_game_discount <- function(df, target_title, known_discount) {
  df |>
    mutate(
      discount_pct = ifelse(title == target_title, known_discount, discount_pct),
      original_price = ifelse(
        title == target_title, 
        current_price / (1 - discount_pct / 100), 
        original_price
      )
    )
}

```

### Validate function
Scaleable Validation Function - identifies rows where the maths is incorrect (we include a tolerance as we saw sales etc. that changed the prices a lot)

```{r}
validate_price_math <- function(df, tolerance = 0.05) {
  df |>
    mutate(
      expected_price = round(original_price * (1 - discount_pct / 100), 2),
      price_difference = abs(current_price - expected_price)
    ) |>
    # Return ONLY the rows that fail the validation check
    filter(price_difference > tolerance)
}
```

## Scrape

### Start server and browser

```{r}
# Start the Selenium server (temp = FALSE keeps it across sessions)
server <- selenium_server(temp = FALSE)
wait_for_server(server)

# Open a browser session (defaults to Firefox)
driver <- SeleniumSession$new()

```

### Calling the functions

```{r}
# 1. Define URLs
relevance_link <- "https://store.steampowered.com/search/?specials=1"
reviews_link <- "https://store.steampowered.com/search/?sort_by=Reviews_DESC&supportedlang=english&specials=1&ndl=1"
release_date_link<- "https://store.steampowered.com/search/?sort_by=Released_DESC&supportedlang=english&specials=1&ndl=1"

# 2. Harvest pages
raw_relevance <- harvest_steam_page(relevance_link, target_games = 500) |> 
  mutate(sort_method = "Relevance")

raw_reviews <- harvest_steam_page(reviews_link, target_games = 500) |>
  mutate(sort_method = "User Reviews")

raw_release_date <- harvest_steam_page(release_date_link, target_games = 500) |>
  mutate(sort_method = "Release date")

# 3. Clean results
clean_relevance <- raw_relevance |>
  clean_steam_prices() |>
  impute_game_discount(target_title = "NBA 2K26", known_discount = 71)

clean_reviews <- raw_reviews |>
  clean_steam_prices()

clean_release_date <- raw_release_date |>
  clean_steam_prices()

View(clean_relevance)
View(clean_reviews)
View(clean_release_date)

```

### Close browser and stop server

```{r}
driver$close()
server$kill()
```

## Results

### Check inconsistencies / Missing Values

```{r}
clean_relevance |>
  validate_price_math(tolerance = 0.05)

clean_reviews |>
  validate_price_math(tolerance = 0.05)

clean_release_date |>
  validate_price_math(tolerance = 0.05)
```

```{r}
plot_intro(clean_relevance)
plot_intro(clean_reviews)
plot_intro(clean_release_date)
```

### Write csv

```{r}
write_csv(clean_relevance, file = "clean_relevance.csv")
write_csv(clean_reviews, file = "clean_reviews.csv")
write_csv(clean_release_date, file = "clean_release_date.csv")
```

# API Integration

We store the API key in a local .txt file to ensure it's not pushed to the GitHub repository, following security best practices. The .gitignore file has been configured to exclude 'api_key.txt'.

```{r}
if (!file.exists("api_key.txt")) {
  stop("REPRODUCIBILITY ERROR: 'api_key.txt' not found. Please create this file 
       in the root directory and paste your RAWG API key to proceed.")
}

key <- readLines("api_key.txt", warn = FALSE)


# Loading dataset generated in the Scraping phase
steam_data <- read.csv("clean_reviews.csv")

# Check column types and distribution
summary(steam_data)
```

Although we performed an initial cleaning of names and prices during the scraping above, Steam titles remained "noisy" for external API integration.

Many titles still contained trademarks (™), regional subtitles, or localised translations (e.g., "Valiant Hearts / Soldats Inconnus") that prevented exact matches in the RAWG database. Without this secondary normalization layer, our API match would have been significantly lower, leading to a data loss of nearly 60%. This layer ensures inter-operability between two different data sources (and proves we love R!).

## Data Diagnostics & String Normalisation

```{r}
# Find special characters and symbols that cause errors
symbol_count <- sum(grepl("™|®|\\(|\\)", steam_data$title))
print(paste("Number of titles requiring advanced normalization:", symbol_count))

# Regex to tidy
steam_data <- steam_data %>%
  mutate(title_search = title %>%
    gsub("™|®", "", .) %>%
    gsub("\\(.*\\)", "", .) %>%
    gsub("/.*", "", .) %>%
    gsub(":.*", "", .) %>%
    gsub(" - .*", "", .) %>%
    trimws()
  )

```

## API Enrichment Function
This function connects to the RAWG API to retrieve the Metacritic score and Genre.

We decided to implement a "Best Match" kind of logic to avoid DLC or Soundtracks without scores (there were a few strange matches).

```{r}
get_metacritic <- function(game_name, api_key) {
  
  # remove non-alphanumeric characters
  query_clean <- gsub("[^[:alnum:] ]", " ", game_name)
  query_clean <- trimws(gsub("\\s+", " ", query_clean))
  query_url <- URLencode(query_clean)
  
  url <- paste0("https://api.rawg.io/api/games?key=", api_key, "&search=", query_url)
  
  tryCatch({
    resp <- request(url) %>% req_perform()
    data <- resp %>% resp_body_json(simplifyVector = TRUE)
    
    if (length(data$results) > 0) {
      res <- data$results
      # Check if the metacritic column exists and filter for the first entry with a valid score
      if ("metacritic" %in% names(res)) {
        valid_match <- res[!is.na(res$metacritic), ]
        if (nrow(valid_match) > 0) {
          return(data.frame(
            metacritic = valid_match$metacritic[1],
            genre = valid_match$genres[[1]]$name[1],
            stringsAsFactors = FALSE
          ))
        }
      }
    }
    # Return NA if no match with a score is found
    return(data.frame(metacritic = NA, genre = "Unknown", stringsAsFactors = FALSE))
  }, error = function(e) return(data.frame(metacritic = NA, genre = "Unknown")))
}
```

## Execution and Data Quality Audit

```{r}
print("Starting API Enrichment... This may take a few minutes for 500 records.")

# Iterating through the cleaned titles using lapply
api_list <- lapply(steam_data$title_search, get_metacritic, api_key = key)

# Put the results into a data frame
api_results <- do.call(rbind, api_list)

# Merging frame with the original Steam data set
final_table <- cbind(steam_data, api_results)

# Check the number of successfully retrieved scores
na_count <- sum(is.na(final_table$metacritic))
success_rate <- round((1 - (na_count / nrow(final_table))) * 100, 2)

print(paste("Data Enrichment Success Rate:", success_rate, "%"))
print(paste("Total titles with Metacritic scores retrieved:", nrow(final_table) - na_count))
```

## Export Final Dataset

```{r}
# Exporting the enriched dataset for visualization and reporting
write.csv(final_table, "final_data_steam_rawg.csv", row.names = FALSE)
```

## Data Visualisation and Statistical Analysis

### Visualisation

```{r}
# Filtering data to include only games with a valid Metacritic score
df_plot <- final_table %>% filter(!is.na(metacritic))

# 1. Statistical Correlation
# We calculate the Pearson correlation to see if higher discounts relate to lower quality.
correlation_value <- cor(df_plot$discount_pct, df_plot$metacritic, use = "complete.obs")
print(paste("Pearson Correlation (Quality vs Discount):", round(correlation_value, 3)))

# 2. Scatter Plot: Quality vs. Discount
# This graph helps identify if "hidden gems" (high score, high discount) exist.
ggplot(df_plot, aes(x = discount_pct, y = metacritic)) +
  geom_point(aes(color = genre), size = 3, alpha = 0.6) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Game Quality vs. Discount Percentage",
    x = "Discount (%)",
    y = "Metacritic Score",
    color = "Genre"
  ) +
  facet_wrap(~genre)

# 3. Bar Chart: Average Quality by Genre
# Useful for identifying which genres currently have the best-rated deals.
df_genre <- df_plot %>% 
  group_by(genre) %>% 
  summarise(avg_score = mean(metacritic), count = n()) %>%
  filter(count > 2) %>% # Filtering small samples for better reliability
  arrange(desc(avg_score))

ggplot(df_genre, aes(x = reorder(genre, avg_score), y = avg_score, fill = genre)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Average Metacritic Score by Genre",
    x = "Genre",
    y = "Mean Score"
  )

# 4. Histogram: Discount Distribution
# A general overview of the Steam sales landscape.
ggplot(final_table, aes(x = discount_pct)) +
  geom_histogram(fill = "#5DADE2", bins = 20, color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Discounts across Top 500 Games",
    x = "Discount (%)",
    y = "Number of Games"
  )
```

### Quality vs Value analysis
This is our final visualisation, showing value and discount by genre, with an indication of what that means as a consumer.

```{r}
# 1. Cleaned Data Analysis
analisis_genres <- final_table %>%
  filter(!is.na(metacritic)) %>%
  group_by(genre) %>%
  summarise(
    total_games = n(),
    avg_metacritic = mean(metacritic),
    avg_discount = mean(discount_pct)
  ) %>%
  filter(total_games > 5) %>%
  mutate(value_score = avg_metacritic * (avg_discount / 100)) %>% 
  arrange(desc(value_score))

# 2. Calculate dynamic center points
mid_x <- mean(analisis_genres$avg_discount, na.rm = TRUE)
mid_y <- mean(analisis_genres$avg_metacritic, na.rm = TRUE)

# 3. Final Quadrant Plot
ggplot(analisis_genres, aes(x = avg_discount, y = avg_metacritic)) +
  # Background Colors
  annotate("rect", xmin = mid_x, xmax = Inf, ymin = mid_y, ymax = Inf, fill = "green", alpha = 0.07) + 
  annotate("rect", xmin = -Inf, xmax = mid_x, ymin = mid_y, ymax = Inf, fill = "blue", alpha = 0.07) +  
  annotate("rect", xmin = mid_x, xmax = Inf, ymin = -Inf, ymax = mid_y, fill = "orange", alpha = 0.07) + 
  annotate("rect", xmin = -Inf, xmax = mid_x, ymin = -Inf, ymax = mid_y, fill = "red", alpha = 0.07) + 

  # Quadrant Labels
  annotate("text", x = Inf, y = Inf, label = "BUY NOW\n(High Rating, High Discount)", 
           hjust = 1.1, vjust = 1.5, fontface = "bold", color = "darkgreen", size = 4) +
  annotate("text", x = -Inf, y = Inf, label = "IF YOU LOVE THEM\n(High Rating, Low Discount)", 
           hjust = -0.1, vjust = 1.5, fontface = "bold", color = "darkblue", size = 4) +
  annotate("text", x = Inf, y = -Inf, label = "STOCKING FILLERS\n(Low Rating, High Discount)", 
           hjust = 1.1, vjust = -0.5, fontface = "bold", color = "darkorange", size = 4) +
  annotate("text", x = -Inf, y = -Inf, label = "SKIP\n(Low Rating, Low Discount)", 
           hjust = -0.1, vjust = -0.5, fontface = "bold", color = "darkred", size = 4) +

  # Reference Lines
  geom_vline(xintercept = mid_x, linetype = "dashed", color = "gray70") +
  geom_hline(yintercept = mid_y, linetype = "dashed", color = "gray70") +

  # Data Layers (Points and Genre Names)
  geom_point(aes(size = total_games), color = "black", alpha = 0.4) +
  geom_text(aes(label = genre), vjust = -1.2, size = 3.5, fontface = "bold", check_overlap = TRUE) +

  # Styling & Scales
  scale_size_continuous(range = c(4, 18), name = "Number of Games") +
  labs(
    title = "Game Genre Value Analysis",
    subtitle = paste("Quadrants centered on Avg Discount (", round(mid_x, 1), "%) and Score (", round(mid_y, 1), ")"),
    x = "Average Discount (%)",
    y = "Average Metacritic Score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "bottom"
  )
```

